---
title: 
author:
date: 
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Part 2: Modelling a building using CTSM-R
## 2a: 2-state model of a single room

We will estimate 

```{r}
# Load data
if (Sys.info()[7] == "davidipsen")
  {path <- "~/Documents/DTU/3. Semester (MSc)/Advanced Time Series/Assignments/3-Assignment/CompEx3_E18/"
} else {path <- "CompEx3_E18/"}
load(paste0(path,"Exercise3.RData"))
D = AllDat
head(D)
tail(D)
```


Visualisations
```{r}
plot(D$Ta * D$Gv, D$yTi1)
```

Let's try expanding the model.
1. Note how the effect of the temperature in room 3 and 4 affects only room 1 through the medium of room 2. In other words: $Y_1$ is *conditionally independent* of $Y_3$ and $Y_4$ when $Y_2$ is known.
---- disadvantage: Time delay for a change in T2 before it enters T1 -> enter Tm before -> Make Tm also a function of T2?
==
Men hey, er det ikke bare via Thermal Mass (Tm) at room1 bliver påvirket?
2. Dvs. dTm = ...Org.... + 1/R_21*(T2-T1x)



