---
title: 
author:
date: 
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Part 2: Modelling a building using CTSM-R
## 2a: 2-state model of a single room

We will estimate the following model where the impact of the measured radiation $G_v$ is scaled relative the sun's angle through the window. Since we do not have access to this information, a non-parametric fit will be done using B-splines.

$$\begin{aligned} d T_{i} &=\frac{1}{C_{i}}\left(\frac{1}{R_{i a}}\left(T_{a}-T_{i}\right)+\frac{1}{R_{i m}}\left(T_{m}-T_{i}\right)+\Phi+\left(\sum_{k=1}^{N} a_{k} b s_{k}(t)\right) G_{v}\right) d t+\sigma_{1} d w_{1} \\ d T_{m} &=\frac{1}{C_{m}}\left(\frac{1}{R_{i m}}\left(T_{i}-T_{m}\right)\right) d t+\sigma_{2} d w_{2} . \\ y T_{i} &=T_{i}+e_{1}, \end{aligned}$$


```{r, results="hide"}
#install.packages("ctsmr", repo = "http://ctsm.info/repo/dev")
#install.packages("pkgbuild")

# For git pushing
## git push https://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@github.com/davidripsen/3-Assignment.git

library(ctsmr)
library(splines)
source("CompEx3_E18/sdeTiTm.R")
# Load data
if (Sys.info()[7] == "davidipsen")
  {path <- "~/Documents/DTU/3. Semester (MSc)/Advanced Time Series/Assignments/3-Assignment/CompEx3_E18/"
} else {path <- "CompEx3_E18/"}
load(paste0(path,"Exercise3.RData"))
#AllDat

########## Initial model ############
  fit1 <- sdeTiTm(AllDat,AllDat$yTi1,AllDat$Ph1) # Original model
  
  summary(fit1,extended=TRUE)
  fit1$loglik
  
  Hour <- as.numeric(strftime(AllDat$date, format="%H"))
  
  Pred <- predict(fit1)
  #plot(Pred[[1]]$state$pred$Ti - AllDat$yTi1 ~ Hour)
  
  # Fit only splines for radiation hours
  #plot(AllDat$Gv ~ Hour) #
  
  
  idx <- (Hour>8 & Hour < 23) # It is impossible to fit a window area for the hours without any sun, so we limit the window area estimation to the hours with sun.
  bs = bs(Hour[idx],df=5,intercept=TRUE) # Dvs. 4 knots / 5 basis splines
  
  # What does the splines look like?
  #plot(bs[14:27,1],type='l')
  #lines(bs[ 14:27,2])
  #lines(bs[ 14:27,3])
  #lines(bs[ 14:27,4])
  #lines(bs[ 14:27,5])
  
  bs1 <- bs2 <- bs3 <- bs4 <- bs5 <- bs6 <- numeric(dim(AllDat)[1])
  
  bs1[idx] = bs[,1]
  bs2[idx] = bs[,2]
  bs3[idx] = bs[,3]
  bs4[idx] = bs[,4]
  bs5[idx] = bs[,5]
  
  AllDat$bs1 = bs1
  AllDat$bs2 = bs2
  AllDat$bs3 = bs3
  AllDat$bs4 = bs4
  AllDat$bs5 = bs5


### IMPLEMENT THE MENTIONED MODEL ###
source(paste0(path,"sdeTiTmAv.R"))
fit2 <- sdeTiTmAv(AllDat,AllDat$yTi1,AllDat$Ph1)
```

Let's compare the two models

```{r}
sprintf('Model 1: logL = %f', fit1$loglik)
sprintf('Model 2: logL = %f', fit2$loglik)
```

I.e. we see a very large improvement in likelihood (for only 4 extra parameters).

```{r}
summary(fit2, extended=T)
```

We see that all parameters are significant, except for a2. For completeness, a2 is kept in the model. Let's visualize the spline-fit


```{r}
plot(9:22, bs[14:27,1]*fit2$xm[3]+bs[14:27,2]*fit2$xm[4]+bs[14:27,3]*fit2$xm[5]+bs[14:27,4]*fit2$xm[6]+bs[14:27,5]*fit2$xm[7],type='l')
```

Remember, that the above fit is not the actual input radiation to the room, but it is a *weighting* of the actual radiation. Apparently, the model would like to have more emphasis on the radiation when the sun is low.






# 2b: Improving the single-room model
In this part, I will try to expand the model.
First, note how the effect of the temperature in room 3 and 4 affects only room 1 through the medium of room 2. In other words: $T_1$ is *conditionally independent* of $T_3$ and $T_4$ when $T_2$ is known.

Initially, I added the temperature of the neighboring room ($T_2$) as an additional state and then had it impact $dT_1$. However, I realized that the temperature of room 2 only affects room 1 through *the wall* as a medium, i.e. the thermal mass of room 1. And since we have direct measurements of $T2$ and only focus on modelling the single-room ($T1$), I add it as input to the model with the term: $$\frac{1}{R_{2,1}}(T_2-T_1)$$
thereby letting the change in thermal mass of room 1 ($dT_m$) being proportional to the difference in temperature between the two rooms.

The model is then

$$\begin{aligned} d T_{1} &=\frac{1}{C_{1}}\left(\frac{1}{R_{i a}}\left(T_{a}-T_{1}\right)+\frac{1}{R_{i m}}\left(T_{m}-T_{1}\right)+\Phi+\left(\sum_{k=1}^{N} a_{k} b s_{k}(t)\right) G_{v}\right) d t+\sigma_{1} d w_{1} \\
d T_{m} &=\frac{1}{C_{m}}\left(\frac{1}{R_{i m}}\left(T_{1}-T_{m}\right) + \frac{1}{R_{2,1}}(T_2-T_1)\right) d t+\sigma_{2} d w_{2} . \\ y T_{1} &=T_{1}+e_{1}, \end{aligned}$$

Let's estimate it
```{r, results="hide"}
source(paste0(path,"2b-sdeT1T2TmAv.R"))
fit3 <- sdeT1T2TmAv(AllDat,AllDat$yTi1,AllDat$Ph1)
```

Let's have a look at the model.
```{r}
sprintf('Model 3: logL = %f', fit3$loglik)
sprintf('Likelihood ratio test: p-value = %f', 1-pchisq(abs(2*(fit2$loglik - fit3$loglik)),1))
summary(fit3, extended=T)
```

Again, a large improvement in likelihood is seen (at the cost of 1 extra parameter). Since the two models are nested, at likehood ratio test is performed and shows a very strong signficant difference. Again, all parameters are significant, except for the observation error which might in fact have 0 as the true value, however we do see some strong correlations in the coefficients (!).

Now expanding further on the model, I see that the radiator is placed close to the window in room 1, i.e. some of the heating might go straight out of the window. Additionally, I note that the North Heating Circuit and South Heating Circuit are piped together before entering the building, so the actual heating in room 1 is a mix of the two (however, the two are higly correlated with $\hat{\rho} = 0.91$). I.e. denote = $\Phi = (\Phi_1+\Phi_2)/2$ and change then the $\Phi$ term in $dT_1$ to the term $(1-c) \Phi - c\Phi \frac{T_1-T_a}{T_1}$, where $c \in [0,1]$. This makes a proportion $c$ of the heating load escape directly through the window relative to the temperature ratio between inside and outside\footnote{The meaning of $T_a$ is not explicit from the assignment. From plots I've concluded to assume it to be the outside temperature.}. Note that using this formulation, the hypothesis H0: $c=0$  is of interest (meaning no loss of heating) and when the outdoor and indoor temperature is the same, the second term cancels (meaning no loss of heating).

The model then becomes
$$\begin{aligned} d T_{1} &=\frac{1}{C_{1}}\left(\frac{1}{R_{i a}}\left(T_{a}-T_{1}\right)+\frac{1}{R_{i m}}\left(T_{m}-T_{1}\right)+ ((1-c) \Phi - c\Phi \frac{T_1-T_a}{T_1}) +\left(\sum_{k=1}^{N} a_{k} b s_{k}(t)\right) G_{v}\right) d t+\sigma_{1} d w_{1} \\
d T_{m} &=\frac{1}{C_{m}}\left(\frac{1}{R_{i m}}\left(T_{1}-T_{m}\right) + \frac{1}{R_{2,1}}(T_2-T_1)\right) d t+\sigma_{2} d w_{2} . \\ y T_{1} &=T_{1}+e_{1}, \end{aligned}$$

Where $\Phi = (\Phi_1+\Phi_2)/2$. 

```{r, results="hide"}
#plot(AllDat$date, AllDat$Ta, type='l')
#plot(AllDat$date, (AllDat$Ph1 + AllDat$Ph2)/2, type='l')
#plot(AllDat$Ph1, AllDat$Ph2); cor(AllDat$Ph1, AllDat$Ph2)
source(paste0(path,"2b-2-sdeT1T2TmAv.R"))
fit4 <- sdeT1T2TmAv(AllDat,AllDat$yTi1)
```

Let's review the quality of the fit

```{r}
fit4$loglik
summary(fit4, extended=T)

sprintf('Likelihood ratio test from previous model: p-value = %f', 1-pchisq(-2*(fit3$loglik - fit4$loglik),1))
```











# 3 Making a multi-room model

$$
\begin{aligned} d T_{1} &=\frac{1}{C_{1}}\left(\frac{1}{R_{i a}}\left(T_{a}-T_{1}\right)+\frac{1}{R_{i m}}\left(T_{m}-T_{1}\right)+\Phi+ A_w G_{v}\right) d t+\sigma_{1} d w_{1} \\

d T_{1m} &=\frac{1}{C_{m}}\left(\frac{1}{R_{i m}}\left(T_{1}-T_{m}\right) + \frac{1}{R_{2,1}}(T_2-T_1)\right) d t+\sigma_{2} d w_{2} . \\


d T_{1} &=\frac{1}{C_{1}}\left(\frac{1}{R_{i a}}\left(T_{a}-T_{1}\right)+\frac{1}{R_{i m}}\left(T_{m}-T_{1}\right)+\Phi+ A_w G_{v}\right) d t+\sigma_{1} d w_{1} \\

d T_{1m} &=\frac{1}{C_{m}}\left(\frac{1}{R_{i m}}\left(T_{1}-T_{m}\right) + \frac{1}{R_{2,1}}(T_2-T_1)\right) d t+\sigma_{2} d w_{2} . \\

d T_{1} &=\frac{1}{C_{1}}\left(\frac{1}{R_{i a}}\left(T_{a}-T_{1}\right)+\frac{1}{R_{i m}}\left(T_{m}-T_{1}\right)+\Phi+ A_w G_{v}\right) d t+\sigma_{1} d w_{1} \\

d T_{1m} &=\frac{1}{C_{m}}\left(\frac{1}{R_{i m}}\left(T_{1}-T_{m}\right) + \frac{1}{R_{2,1}}(T_2-T_1)\right) d t+\sigma_{2} d w_{2} . \\


d T_{1} &=\frac{1}{C_{1}}\left(\frac{1}{R_{i a}}\left(T_{a}-T_{1}\right)+\frac{1}{R_{i m}}\left(T_{m}-T_{1}\right)+\Phi+ A_w G_{v}\right) d t+\sigma_{1} d w_{1} \\

d T_{1m} &=\frac{1}{C_{m}}\left(\frac{1}{R_{i m}}\left(T_{1}-T_{m}\right) + \frac{1}{R_{2,1}}(T_2-T_1)\right) d t+\sigma_{2} d w_{2} . \\

\forall_i y T_{i} &=T_{i}+e_{i} \end{aligned}$$









